---
title: "MERS"
output: html_notebook
---

## Load libraries

```{r}
library(ape)
library(treedater)
library(adephylo)
library(lubridate)
library(ggplot2)
```

## Read sequence data

```{r}
s <- read.dna("MERS_CoV_274_CDS.fna",format="fasta")
s
```

Load dates and traits.

```{r}
snames <- row.names(s)
sdates <- read.table("MERS_CoV_274_CDS.dates.txt",header=F,row.names=NULL,sep="\t")
straits <- read.table("MERS_CoV_274_CDS.traits.txt",header=F,row.names=NULL,sep="\t")
```

Rename sequences to conform with IQTREE specifications.

```{r}
newsnames <- gsub("/","_",snames,fixed=TRUE)
newsnames <- gsub("|","_",newsnames,fixed=TRUE)
newsnames <- gsub("_","_",newsnames,fixed=TRUE)
news <- s
row.names(news) <- newsnames
write.dna(news,"MERS_CoV_274_CDS_renamed.fasta",format="fasta",nbcol=-1,colsep="")
```

Run IQTREE.

```{bash}
iqtree -s MERS_CoV_274_CDS_renamed.fasta -spp partitions.nex -pre MERS_CoV_274_CDS_renamed -allnni -nt 12
```

## Treedater analysis

```{r}
treefile <- "MERS_CoV_274_CDS_renamed.treefile"
seqlen <- 29364
tre <- unroot(read.tree(treefile))
tipnames <- tre$tip.label
```

```{r}
idx <- match(tipnames,newsnames)
oldnames <- snames[idx]
tipdates <- sdates[match(oldnames,sdates[,1]),2]
names(tipdates) <- tipnames
```

```{r}
treed <- dater(tre, tipdates, s=seqlen, maxit = 100, minblen=1./365, abstol = .001, quiet = TRUE, temporalConstraints=TRUE, numStart=2, searchRoot=10, strictClock=FALSE, ncpu=10)
treed
```

## MCC from Dudas et al.

```{r}
mcc <- read.nexus("MERS_274_sCoal.combinedTyped.mcc.tree")
mcc
```

```{r}
mccidx <- match(mcc$tip.label,sdates[,1])
mccdates <- sdates[mccidx,2]
max(mccdates)
```

This is already a time-calibrated tree, so we simply need to work out the depth in order to work out the TMRCA.

```{r}
mcc.dr <- distRoot(mcc)
```

## LSD

Output data for LSD.

```{r}
df <- data.frame(c(length(tipdates),names(tipdates)),c("",tipdates))
write.table(df,"MERS_CoV_274_CDS.td",col.names=FALSE,row.names=FALSE,quote=FALSE,sep="\t")
```

Run LSD.

```{bash}
lsd -i MERS_CoV_274_CDS_renamed.treefile -d MERS_CoV_274_CDS.td -o MERS_CoV_274_CDS.lsd -r a -v 2 -b 10 -s 29364 -c -f 1000
```

Read in tree and calculate distances to root.

```{r}
lsd <- read.tree("MERS_CoV_274_CDS.lsd.date.newick")
```

```{r}
lsd.dr <- distRoot(lsd)
```

## Comparison TMRCA

```{r}
date_decimal(max(mccdates)-max(mcc.dr))
```

```{r}
date_decimal(treed$timeOfMRCA)
```

```{r}
date_decimal(max(mccdates)-max(lsd.dr))
```

## Comparison LTT

```{r}
ltt <- function(tre,mrsd){
  lineages <- ltt.plot.coords(tre,backward=TRUE)
  lineages <- as.data.frame(lineages)
  names(lineages) <- c("Time","Lineages")
  lineages$Time <- mrsd+lineages$Time
  lineages
}
```

```{r}
treed.ltt <- ltt(treed,decimal_date(ymd("2015-09-17")))
mcc.ltt <- ltt(mcc,decimal_date(ymd("2015-09-17")))
lsd.ltt <- ltt(lsd,decimal_date(ymd("2015-09-17")))
```


```{r}
p <- ggplot()+geom_line(data=treed.ltt,aes(x=Time,y=Lineages,color="treedater"))+geom_line(data=mcc.ltt,aes(x=Time,y=Lineages,color="BRMC"))+geom_line(data=lsd.ltt,aes(x=Time,y=Lineages,color="QPD"))+scale_alpha_manual(name="Model", values=c("treedater"="red","BRMC"="purple","QPD"="green"))+scale_x_continuous(limits=c(2008,2016))
p
```


